### 2021-12-09 preliminary_analysis.Rmd
more data

---
title: "preliminary_regression"
author: "Deniz Tokmakoglu"
date: "12/7/2021"
output: html_document
---

```{r setup, include=FALSE}
library("tidyverse")
library("knitr")
library("stargazer")
library(readstata13)
library(ggplot2)
library("stringr")
library(data.table)
library("reticulate")
library(fixest)
library(sandwich)
library(tidyverse)
library(lubridate)
library(jtools)
library(rdd)
library(ggplot2)
library(WeightIt)
library(DescTools)
library(knitr)
library(extrafont)
library(splines)
library(lubridate)
library(stargazer)
library(dplyr)
library(data.table)
library(janitor)
library(gcookbook)
library(plm)
library(lfe)
library(modelsummary)
library(formattable)
library(lmtest)
```

```{r}
election_all <- read_csv("clean_data/elections_all.csv")
```

# Initial Regression
```{r}
election_all <- election_all %>% 
  mutate(crisis = ifelse(is.na(`crisis type`), 0, 1),
         percent_turnout = ifelse(percent_turnout < 1, 100 * percent_turnout, 
                                  percent_turnout))

reg1 <- lm(percent_turnout ~ crisis, data = election_all)

stargazer(reg1, type = "text")
```

# By Crisis Type

```{r}
reg1 <- lm(percent_turnout ~ `crisis type_civil unrest` + `crisis type_coup` +
             `crisis type_economic` + `crisis type_refugee` + `crisis type_regufee` + 
             `crisis type_soft coup` + `crisis type_terrorism` + `crisis type_war`
, data = election_all)
```