### 2021-12-11 paper/main.tex
paper_changes

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{float}
\usepackage{booktabs,dcolumn}
\usepackage{indentfirst}
\usepackage{hyperref}
\usepackage{appendixnumberbeamer}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{outlines} % new package, delete if getting package errors
\usepackage[scale=2]{ccicons}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepgfplotslibrary{dateplot}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{xspace}
\usepackage{tikz}
\usepackage{newtxtext}
\usepackage{titlesec}
\usepackage{lscape}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\newcolumntype{d}{D{.}{.}{2.5}}           % alignment on decimal marker
\newcommand\mc[1]{\multicolumn{1}{c}{#1}}
\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-1in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Turkish Elections and Catastrophic Events}
    {
}

  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title}
\end{center}
  \medskip
} \fi

\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
\label{sec:intro}

A year ago, Turkey celebrated a century of being a democracy. Now, only 2 years away from the same milestone for being a republic, it is inevitable to think and analyse the trends of Turkish democracy. 
It is reasonable to say, Turkey, which resides possibly one of the lesser stable regions in the world; suffered a lot during this period, while trying to establish their democratic traditions. Having survived 3 coups, 3 "soft" coups and numerous economic, political and civil crises, Turkey still maintained regular and somewhat fair elections without interruption, unlike many other countries in a similar political landscape. The goal of this paper is to materialise this continuity and uncover the impact of catastrophic events in Turkey to voting behavior and results.

\section{Background}

\subsection{Brief Summary of Turkish History}

After the end of World War I, like many empires in Europe; Ottoman Empire crumbled. In it's heartland, Turkey emerged as a new republic in 1923 after a long period of successive independence war against the colonial powers.

Founded as a parliamentary republic, the first 17 years of the country has been defined by a single party period in which the founding party Republican People's Party, (Turkish abbreviation CHP will be used thereafter) was in power. The party is defined by having pro-European, Liberal and Secular values. Only in 1950, Turkey had it's truly first election with emergence of Right wing parties. Between 1950 - 2018, Turkey remained as a parliamentary republic with periodic municipal and parliamentary elections.

In 2018, with a 51\% referendum result, Turkey decided to transition into a presidential republic where the president and the Parliament is elected at the same time, using different ballots. For this project's scope Turkey's only 2 presidential elections were excluded from research.

Inheriting many problems of it's predecessor, since it's foundation Turkey needed to deal with both regional and domestic problems. Composed of many different ethnic backgrounds such as Turkish, Kurdish, Armenian, Zaza etc. Turkey had to actively fight to maintain it's regional integrity. Moreover, as the gateway of Middle East to Europe, Turkey suffered from many regional events taking place outside of it's borders. One great example of this is the refugee influx to Turkey after ISIS war in Syria in 2011.

\subsection{Summary of Catastrophic Events}

The research consists of 4 main catastrophic events, which will be used to analyze voter behaviour.

\subsubsection{1999 Marmara Earthquake}
On August 17, 1999 the Marmara Region in Turkey was devastated from a 7.5 magnitude earthquake that lasted for several days. Most of the buildings in the epicenter of the earthquake were either severely damaged or collapsed by the end of the event. Approximately 18,000 people were killed. It was a showcase of Turkey's unpreparedness for earthquakes, a country which sits on various earthquake lines.

\begin{subfigure}[t]{.49\textwidth}
    \includegraphics[scale=0.15]{paper/figures/earthquake_marmara2.jpeg}
	\caption{Marmara Earthquake}
	\end{subfigure}
	\begin{subfigure}[t]{.49\textwidth}
    \includegraphics[scale=0.3]{paper/figures/marmara_affected.jpeg}
	\caption{Regions Affected}
\end{subfigure}

\subsubsection{1974 Cyprus Peace Operation}

The island of Cyprus, which used to be ruled entirely by Republic of Cyprus lies in the Eastern Mediterranean, approximately 70km away from the Turkish mainland. The island is composed of Cypriot Greeks and Cypriot Turks. While the latter is the minority, the founding principles of the Republic of Cyprus ensured equal representation for both, and assigned Turkey, Greece and the UK as guarantor states. All had the ability to intervene if the ethnic groups they represent were in danger. The tension in the island was ever increasing with Turkish minority under attack by Greek Nationalist organizations but the tipping point was in 1974; when a pro Greek unification president got elected in the island.

Turkey decided to intervene, causing relocation of many Greek and Turkish locals and separation of the island into 2. The north declared it's independence as Turkish Republic of Northern Cyprus; which is only recognized by Turkey.

\subsubsection{2011 Refugee Crisis}

In 2011, a civil war started in Syria after a set of peaceful student demonstrations caused the government's crackdown. 80\% of the Syrians got displaced and the majority of them entered to Turkey via the border, making Turkey the largest refugee holder in the world since 2014.

\subsubsection{1984 Conflict with Kurdish Armed Groups}

Composing one fifth of Turkey's population, ethnic Kurdish people mostly live in the Eastern cities in Turkey (mainly Van, Diyarbakir). The PKK (Kurdish Worker's Party), established by Abdullah Ocalan in 1978, has waged an insurgency since 1984 against Turkish authorities for greater cultural and political rights, primarily with the objective of establishing an independent Kurdish state. The ongoing conflict in the eastern region has resulted in nearly forty thousand deaths. The organization has been labelled as a terrorist organization by Turkey and it's allies.

\section{Data and Limitations}

The data is retrieved retrieved from Turkish Statistical Institute, which is the main source for census and election data in Turkey. It consists municipal and parliamentary election results. It consists of the election year, city, election type, percentage turnout, winning party and the vote share, previous winning party and vote share.

The process of data cleaning and creation can be found in

--

\section{Research Design}

\section{Results}

\section{Conclusion}

\end{document}