### 2021-12-11 preliminary_analysis.Rmd
graphics

---
title: "preliminary_regression"
author: "Deniz Tokmakoglu"
date: "12/7/2021"
output: html_document
---

```{r setup, include=FALSE}
library("tidyverse")
library("knitr")
library("stargazer")
library(readstata13)
library(fastDummies)
library("tidyverse")
library("knitr")
library("stargazer")
library(readstata13)
library(ggplot2)
library("stringr")
library(data.table)
library("reticulate")
library(fixest)
library(sandwich)
library(lmtest)
library(ggplot2)
library("stringr")
library(data.table)
library("reticulate")
library(fixest)
library(sandwich)
library(tidyverse)
library(lubridate)
library(jtools)
library(rdd)
library(ggplot2)
library(WeightIt)
library(DescTools)
library(knitr)
library(extrafont)
library(splines)
library(lubridate)
library(stargazer)
library(dplyr)
library(data.table)
library(janitor)
library(gcookbook)
library(plm)
library(lfe)
library(modelsummary)
library(formattable)
library(lmtest)
```

```{r}
election_all <- read_csv("clean_data/all_data_v_event2.csv")
```

# Initial Regression
```{r}

election_all <- election_all %>% mutate(event_right_after = ifelse(year_int - event_year <= 4, 1, 0))

election_all <- election_all %>% 
  mutate(percent_turnout = ifelse(percent_turnout < 1, 100 * percent_turnout, percent_turnout))

reg1 <- felm(percent_turnout ~ event_right_after, data = election_all)

reg2 <- felm(percent_turnout ~ event_right_after + kurdish + refugee | city + year , data = election_all)

stargazer(reg1, reg2, type = "text")
```

# Event Study

```{r}
election_all2 <- election_all %>% select(city, year_int, election_type,
                                      percent_turnout, event_year, event)

election_all2 <- election_all2 %>% 
  mutate(time_to_treat = ifelse(event_year == 0, 0, round((year_int - event_year) / 4)),
         treat = ifelse(event_year == 0, 0, 1))

election_all2 <- election_all2 %>% filter(time_to_treat <= 10)

mod_twfe = feols(percent_turnout ~ i(time_to_treat, treat, ref = -1) ## Our key interaction: time × treatment status
		 |city + year_int,                             ## FEs
		 cluster = ~city ,                          ## Clustered SEs
		 data = election_all2)

iplot(mod_twfe, 
      xlab = 'Time to treatment',
      main = 'Event study: Staggered treatment (TWFE)')
```
# Diff and Diff

```{r}
election_all3 <- election_all2 %>% filter(abs(time_to_treat) <= 3)

election_all3 <- election_all3 %>% mutate(post = ifelse(time_to_treat > 0, 1, 0))

election_all3$treat_post = election_all3$treat * election_all3$post

reg3 <- felm(percent_turnout ~ treat_post | city + year_int + election_type, data = election_all3)
stargazer(reg3, type = 'latex',
          title = c("Effects of a Catastrophic Event on Voter Turnout"),
          column.labels = c('Voter Turnout'),
          covariate.labels = c('Treatment Effect'),
          dep.var.labels = c(''))
```
# Swing

```{r}
election_all4 <- election_all %>%select(city, year_int, election_type,
                                      percent_turnout, event_year, event, swing,
                                      incumbent_vote_share, prev_winner_voteshare)

election_all4 <- election_all4 %>% 
  mutate(time_to_treat = ifelse(event_year == 0, 0, 
                                round((year_int - event_year) / 4)),
         treat = ifelse(event_year == 0, 0, 1))

election_all4 <- election_all4 %>% mutate(post = 
                                            ifelse(time_to_treat > 0, 1, 0))

election_all4 <- election_all4 %>% filter(abs(time_to_treat) <= 3)

election_all4$treat_post = election_all4$treat * election_all4$post

reg4 <- felm(swing ~ treat_post | city + year_int + election_type, 
             data = election_all4)
stargazer(reg4, type = 'latex',
          title = c("Effects of a Catastrophic Event on Change in Leadership"),
          column.labels = c('Dummy for Leadership Change'),
          covariate.labels = c('Treatment Effect'),
          dep.var.labels = c('')
          )
```

# Non Swing Vote Share
```{r}

election_all5 <- election_all4 %>% filter(swing == 0)

election_all5$change_vote = election_all5$incumbent_vote_share - 
  election_all5$prev_winner_voteshare

election_all5$id = paste(election_all5$city, election_all5$year_int, election_all5$election_type)

reg5 <- felm(incumbent_vote_share ~ treat_post | city + year_int + election_type , 
             data = election_all5)

stargazer(reg5, type = 'latex',
          title = c("Effects of a Catastrophic Event on Incumbent Vote Share"),
          column.labels = c('Incumbent Vote Share'),
          covariate.labels = c('Treatment Effect'),
          dep.var.labels = c(''),
          notes = c('Only elections with non-swing result are considered.'))
```

#Swing Vote Share

```{r}
election_all6 <- election_all4 %>% filter(swing == 1)

election_all6$change_vote = election_all6$incumbent_vote_share - 
  election_all6$prev_winner_voteshare

reg7 <- felm(change_vote ~ treat_post | city + year_int, 
             data = election_all6)
stargazer(reg7, type = 'text')
```

# Event Spesific

```{r}
dummies <- dummy_cols(election_all4, select_columns = 'event', ignore_na = TRUE)

dummies <- dummies %>% mutate(event_earthquake = replace_na(event_earthquake, 0),
                              `event_kurdish war` = replace_na(`event_kurdish war`, 0),
                              `event_cyprus war` = replace_na(`event_cyprus war`, 0),
                              event_refugee = replace_na(`event_refugee`, 0))

dummies$treat_post_kurdish = dummies$`event_kurdish war` * dummies$post
dummies$event_earthquake = dummies$event_earthquake * dummies$post
dummies$treat_post_cyprus = dummies$`event_cyprus war` * dummies$post
dummies$treat_post_refugee = dummies$event_refugee * dummies$post

dummies$change_vote = dummies$incumbent_vote_share - dummies$prev_winner_voteshare

reg1_1 <- felm(incumbent_vote_share ~ treat_post_kurdish | city + year_int + election_type, 
             data = dummies)
reg1_2 <- felm(incumbent_vote_share ~ event_earthquake | city + year_int+ election_type, 
             data = dummies)
reg1_3 <- felm(incumbent_vote_share ~ treat_post_cyprus | city + year_int+ election_type, 
             data = dummies)
reg1_4 <- felm(incumbent_vote_share ~ treat_post_refugee | city + year_int+ election_type, 
             data = dummies)

stargazer(reg1_1, reg1_2, reg1_3, reg1_4, type = 'latex',
          title = c("Effects of a Catastrophic Event on Incumbent Vote Share"),
          dep.var.labels = c('Incumbent Vote Share'),
          covariate.labels = c('Kurdish Armed Conflict', 'Marmara Earthquake',
                               'Cyprus Peace Operation', 'Syrian Refugees'),
          notes = c('Only elections with non-swing result are considered.'))

reg2_1 <- felm(change_vote ~ treat_post_kurdish | city + year_int, 
             data = dummies)
reg2_2 <- felm(change_vote ~ event_earthquake | city + year_int, 
             data = dummies)
reg2_3 <- felm(change_vote ~ treat_post_cyprus | city + year_int, 
             data = dummies)
reg2_4 <- felm(change_vote ~ treat_post_refugee | city + year_int, 
             data = dummies)

stargazer(reg2_1, reg2_2, reg2_3, reg2_4, type = 'text')

reg3_1 <- felm(swing ~ treat_post_kurdish | city + year_int, 
             data = dummies)
reg3_2 <- felm(swing ~ event_earthquake | city + year_int, 
             data = dummies)
reg3_3 <- felm(swing ~ treat_post_cyprus | city + year_int, 
             data = dummies)
reg3_4 <- felm(swing ~ treat_post_refugee | city + year_int, 
             data = dummies)

stargazer(reg3_1, reg3_2, reg3_3, reg3_4, type = 'text')

dummies_swing <- dummies %>% filter(swing == 1)

reg4_1 <- felm(incumbent_vote_share ~ treat_post_kurdish | city + year_int, 
             data = dummies_swing)
reg4_2 <- felm(incumbent_vote_share ~ event_earthquake | city + year_int, 
             data = dummies_swing)
reg4_3 <- felm(incumbent_vote_share ~ treat_post_cyprus | city + year_int, 
             data = dummies_swing)
reg4_4 <- felm(incumbent_vote_share ~ treat_post_refugee | city + year_int, 
             data = dummies_swing)

stargazer(reg4_1, reg4_2, reg4_3, reg4_4, type = 'text')

reg5_1 <- felm(percent_turnout ~ treat_post_kurdish | city + year_int, 
             data = dummies)
reg5_2 <- felm(percent_turnout ~ event_earthquake | city + year_int, 
             data = dummies)
reg5_3 <- felm(percent_turnout ~ treat_post_cyprus | city + year_int, 
             data = dummies)
reg5_4 <- felm(percent_turnout ~ treat_post_refugee | city + year_int, 
             data = dummies)

stargazer(reg5_1, reg5_2, reg5_3, reg5_4, type = 'latex',
           title = c("Effects of a Catastrophic Event on Voter Turnout"),
           dep.var.labels = c('Voter Turnout'),
           covariate.labels = c('Kurdish Armed Conflict', 'Marmara Earthquake',
                                 'Cyprus Peace Operation', 'Syrian Refugees'),
            notes = c('Only elections with non-swing result are considered.'))
```